---
title: "Peer Assessment 1"
author: "Mohtashim Shaikh"
date: "16 August 2015"
output: html_document
---

### Loading and preprocessing the data

Read the data and convert date column to date format

```{r}
df <- read.csv("activity.csv")
df$date <- as.Date(df$date)
```

### What is mean total number of steps taken per day?

Below is the histogram of steps taken in this activity

```{r}
agg_df <- na.omit(df)
agg_df <- subset(agg_df, agg_df$steps > 0)
agg_by_date <- aggregate(steps ~ date, data = agg_df, FUN = "sum")
hist(agg_by_date$steps, main = "Histogram of steps", xlab = "Steps")
```

Mean of the total number of steps taken per day

```{r}
aggregate(steps ~ date, data = agg_df, FUN = "mean")
```

Median of the total number of steps taken per day

```{r}
aggregate(steps ~ date, data = agg_df, FUN = "median")
```

### What is the average daily activity pattern?

Plotting the time series of 5-minute interval 

```{r}
agg_m <- aggregate(steps ~ interval, data = df, FUN = "mean")
plot(agg_m$interval, agg_m$steps, type="l", main="Time series plot of 5 mins interval", xlab="Intevals", ylab="Average number of steps per day")
```

Interval with maximum number of steps averaged across all days 
```{r}
subset(agg_m, agg_m$steps == max(agg_m$steps))[1]
```

### Imputing missing values

Total number of missing values in the dataset 

```{r}
nrow(df) - nrow(na.omit(df))
```

Take the mean for that 5-minute interval for all days
```{r}
agg_m <- aggregate(steps ~ interval, data = df, FUN = "mean")


```

### Are there differences in activity patterns between weekdays and weekends?

Add a new column "day" with either "Weekend" or "Weekday"

```{r}
df$tmp <- weekdays(as.Date(df$date))
df$day <- ifelse(df$tmp == "Saturday" | df$tmp == "Sunday", "Weekend", "Weekday")
df$tmp <- NULL
```

